What's hot and what's not:\\Windowing developer topic analysis

### 2009-04-10 lda-paper.tex
global topics

@@

\section{Introduction}

@@

\section{Background}

@@

\subsection{LDA, LSI and Semantic Clustering}

@@

\section{Preliminary Case Study}

In our first exploratory pass we wanted to see if LDA could provide
interesting topics extracted from a real system. We took the
repository of MySQL 3.23, extracted the commits, grouped them by 30
day non-overlapping windows, and then applied LDA to each of these
windows. We asked LDA to make 20 topics per window and we then
 examined the top 10 most frequent words in that topic.  We
found there were common words across topic clusters, such as
\emph{diffs}, \emph{annotate} and \emph{history}, these words probably should
be viewed as stop words. There were notable transitional topics such
as in the first window the word \emph{bitkeeper} appears because MySQL
adopted Bitkeeper for their source control system yet in the following
windows there were no mentions of Bitkeeper. \emph{RENAME} also only
appeared once in the first window and never again. Per each topic we
tried to identify the purpose of the topic; to our surprise we found
that even with only a little familiarity with the code base that
naming the topic was straightforward.

@@

\section{Methodology}
\begin{figure*}
  \centering
  \includegraphics[width=0.9\textwidth]{commit-to-topics} 
  \caption{How commits are analyzed and aggregated into Topics and Trends. Commits are first extracted, then abstracted into word counts or word distributions. These word distributions are then given a topic analysis tool like LDA. LDA finds independent word distributions (topics) that these documents are related to.}
  \label{fig:commits}
\end{figure*}

@@

\subsection{Extraction of  Repositories and Documents}

@@

\subsection{Windowed Sets}

@@

\subsection{Apply Topic Analysis to each Window}

@@

%   6. [ ] Cluster Similarity [0/1]
\subsection{Topic Similarity}
%      1. [ ] need the tool
%             grab from the cluster analysis
%   7. [ ] Continuous Cluster Analysis

@@

\subsection{Visualization}
%       1. [ ] need the visualizer

@@

\subsection{Datasets and Tools}

@@

We extracted the repositories of: PosgreSQL, MaxDB and
Firebird.

@@

\section{Results}

@@

%\subsection{MySQL}
%XXX DO MYSQL?

\subsection{PostgreSQL}
%   3.   mysql

@@

\subsection{Firebird}
%   5.   what's that other db?

@@

\subsection{MaxDB 7.500}

@@

The plots we produced of MaxDB 7.500 were unlike those of the other
systems, as there was a period where no real development occurred and
thus there were no topics or trends whatsoever. Using a topic
similarity of $7/10$ we evaluated of MaxDB 7.500, the first period
from June 2004 to January 2005, and the second period from June 2005
to June 2006.

The largest common trend has references to build system files like
\texttt{SYSDD.punix} and \texttt{MONITOR.punix}.  This trend is
partially displayed at the top of the zoomed in compact-trend-view
(Figure \ref{fig:zoomedsmear}) and at the top of the trend-histogram
(Figure \ref{fig:histogram}).  Other tokens mentioned are
\emph{Sutcheck v1.09}, the prefix SUT stands for Storable Unit
Type. Sutcheck would also automate check-ins using a Perforce SCS tool
which was exporting check-ins to CVS.

@@

The query optimizer was a topic which recurred during
MaxDB's development. In our plots, topics which mention optimizer
occur four times yet in the large static plot (Figure
\ref{fig:statictopics}) it is not in any of the topics. A query
optimizer is rather important to an DBMS, but as we have shown it just
does not pop up as a topic on its own. We tried to remove words to see
if we could get optimizer topic out, after removing stop words and
then two of the most common words the global analysis finally had a
topic with optimizer in its topic ten words. This shows that optimizer
was important but it had been obscured, while it would have been
noticed using a more locally relevant approach that we propose.

We noticed that commits which mentioned \emph{PHP} occured two thirds
less frequently than commits which mentioned \emph{Perl}, but
\emph{Perl} related topics appeared in the global static topics for
MaxDB while \emph{PHP} related topics did not.  Our local topic
analysis mentioned \emph{PHP} in 5 different topics, yet only
mentioned \emph{Perl} in 4 different topics. Perhaps the Perl mentions
were not as independent topics as the PHP topics were.

@@

\subsection{Compare with topics over the entire development}

@@

Previous work on topic analysis that employed LSI and LDA typically
extracted a certain number of topics and tracked them over time. This
means that a single, static set of topics was tracked across the whole
development history of a project.

We carried out the same kind of analysis, as shown in
Figure~\ref{fig:statictopics}. We extracted 20 topics  and plotted
the number of messages per month that were related to that topic. We
found was that one topic would often dominate the results, as shown , while
messages related to other topics did not appear very often.
% This technique seems useful if the 20 extracted topics are relevant
% and useful, that said one weakness of this approach is that documents
% for a topic might only occur in one time window, yet we are tracking
% messages that could be related to it over the entire lifetime of the
% project.
This approach seems reasonable if most of the extracted topics are of
broad interest during most of the development process.  However, it
may be that some topics are of strong interest but only briefly; in
such a case, a window-based model gives a much stronger indication of
the fleeting importance of such topics, and can help to put such a
topic into its proper context.

@@

If we approach the difference of static global topic analysis and
windowed topic analysis via common tokens we can see that common token
tend to dominate both results, at least in the case of MaxDB 7.500 our
local topic approach netted us topics which contained these relatively
important and common words: \emph{UNICODE}, \emph{DEC/OSF1},
\emph{ODBC}, \emph{crash}, \emph{SQLCLI}, \emph{SYSDD.cpnix},
\emph{backup}, \emph{select}, \emph{make}, \emph{memory}, \emph{view},
and finally \emph{debug}. None of these tokens were part of the static
topic analysis topics, but they were part of 566 commits (6\% of the
entire system) to MaxDB 7.500. These tokens were part of 87 out of 520
(26 months, 20 topics per month) of our locally generated topics.

@@

\begin{figure}
  \centering
  %\includegraphics[width=0.45\textwidth]{maxdb7500-everything-by-month}
  \includegraphics[width=0.45\textwidth]{month}
  \caption{MaxDB 7.500 Topics Analyzed with Static Topic Analysis, 20
    topics over the entire development history of MaxDB 7.500 (26 months). The
    shade of the topic indicates the number of documents matching that
    topic in that month relative to the number of documents (white is
    most, black is least).}
  \label{fig:statictopics}
\end{figure}

@@

%   1. well i'm stumped
%   2. look at it?
%   3. investigate those revisions and check the proportion etc
%   4. exploratory
%   5.   work it out
%8.   Validity Threats [0/5]
\section{Validity Threats}

@@

\section{Conclusions}
%11.   Start file file:/home/abez/projects/lda-paper/lda-paper.tex

@@

We proposed and demonstrated the application of windowed topic
analysis, that is, topic analysis applied to commit messages during
periods of development. This approach differed from previous work that
would apply topic analysis globally to a project. We showed that many
topics that exist locally are relevant and interesting yet would often
not be included in global topic analysis. We identify recurring topics
with a topic similarity measure which looks for topics which recur and
mutate throughout the development of the software project.

The value of windowed topic analysis was demonstrated on a case study
of three database systems with a particular focus on MaxDB 7.500. We
showed that static topic analysis missed important topics like ODBC
changes.

@@

We presented multiple visualizations that focused on different aspects
of the data: temporality of trends, trend size, and a
compact-trend-view. The compact-trend-view shows much more
information than the views that global analysis could show, it
indicates how focused a period is by the total number of topics as
well shows topics by similarity so one can track trends across time.
Our trend-histogram immediately hi-lights and measures the size
of trends while our trend-time-line view shows how a topic reoccurs
over time. These visualizations help us get an over feeling about the
common and unique topics that developers focus on. If implemented
properly a user can easily zoom and query a topic.

\subsection{ Future Work}

One avenue of future work we think is quite important is automatic
cluster labelling. Given a word distribution we should be able to
automatically select a word or term that best describes that distribution.
Another avenue we wish to investigate the is value of applying
static community-generated software taxonomies to the naming or
grouping of clusters

>>>>

\newcommand{\shrinkit}{\vspace*{-.3em}}

@@

\shrinkit
\section{Introduction}
\shrinkit

@@

\shrinkit
\section{Background}
\shrinkit

@@

\shrinkit
\subsection{LDA, LSI and Semantic Clustering}
\shrinkit

@@

\shrinkit
\section{Preliminary Case Study}
\shrinkit

In our first exploratory pass we wanted to see if LDA could provide
interesting topics extracted from a real system. We took the
repository of MySQL 3.23, extracted the commits, grouped them by 30
day non-overlapping windows, and then applied LDA to each of these
windows. We asked LDA to make 20 topics per window and we then
 examined the top 10 most frequent words in that topic.  We
found there were common words across topic clusters, such as
\emph{diffs}, \emph{annotate} and \emph{history}, these words probably should
be viewed as stop words. There were notable transitional topics such
as in the first window the word \emph{BitKeeper} appears because MySQL
adopted Bitkeeper for their source control system yet in the following
windows there were no mentions of Bitkeeper. \emph{RENAME} also only
appeared once in the first window and never again. Per each topic we
tried to identify the purpose of the topic; to our surprise we found
that even with only a little familiarity with the code base that
naming the topic was straightforward.

@@

\shrinkit
\section{Methodology}
\shrinkit
\begin{figure*}
  \centering
  \includegraphics[width=0.9\textwidth]{commit-to-topics} 
  \caption{How commits are analyzed and aggregated into Topics and Trends. Commits are first extracted, then abstracted into word counts or word distributions. These word distributions are then given a topic analysis tool like LDA. LDA finds independent word distributions (topics) that these documents are related to.}
  \label{fig:commits}
\end{figure*}

@@

\shrinkit
\subsection{Extraction of  Repositories and Documents}
\shrinkit

@@

\shrinkit
\subsection{Windowed Sets}
\shrinkit

@@

\shrinkit
\subsection{Apply Topic Analysis to each Window}
\shrinkit

@@

%   6. [ ] Cluster Similarity [0/1]
\shrinkit
\subsection{Topic Similarity}
\shrinkit
%      1. [ ] need the tool
%             grab from the cluster analysis
%   7. [ ] Continuous Cluster Analysis

@@

\shrinkit
\subsection{Visualization}
\shrinkit
%       1. [ ] need the visualizer

@@

\shrinkit
\subsection{Datasets and Tools}
\shrinkit

@@

We extracted the repositories of: PostgreSQL, MaxDB and
Firebird.

@@

\shrinkit
\section{Results}
\shrinkit

@@

%\subsection{MySQL}
%XXX DO MYSQL?
\shrinkit
\subsection{PostgreSQL}
\shrinkit
%   3.   mysql

@@

\shrinkit
\subsection{Firebird}
\shrinkit
%   5.   what's that other db?

@@

\shrinkit
\subsection{MaxDB 7.500}
\shrinkit

@@

The plots we produced of MaxDB 7.500 were unlike those of the other
systems, as there was a period where no real development occurred and
thus there were no topics or trends whatsoever (see the gap in Figure
\ref{fig:topicsmear}). Using a topic similarity of $7/10$ we evaluated
of MaxDB 7.500, the first period from June 2004 to January 2005, and
the second period from June 2005 to June 2006.

The largest common trend has references to build system files like
\texttt{SYSDD.punix} and \texttt{MONITOR.punix}.  This trend is
partially displayed at the top of the zoomed in compact-trend-view
(Figure \ref{fig:zoomedsmear}) and at the top of the trend-histogram
(Figure \ref{fig:histogram}).  Other tokens mentioned are
\emph{Sutcheck v1.09} (the prefix SUT stands for Storable Unit
Type), a tool that would also automate check-ins using a Perforce SCS tool
which was exporting check-ins to CVS.

@@

The query optimizer was a topic which recurred during MaxDB's
development. In our plots, topics which mention optimizer occur four
times yet in the global-trend-view (Figure \ref{fig:statictopics}) it
is not in any of the topics. A query optimizer is rather important to
an DBMS, but as we have shown it just does not pop up as a topic on
its own. We tried to remove words to see if we could get an topic
optimizer topic. After removing stop words and then two of the most
common words the global analysis finally had a topic with optimizer in
its topic ten words. This shows that optimizer was important but it
had been obscured, while it would have been noticed using a more
locally relevant approach that we propose.

We noticed that commits which mentioned \emph{PHP} occurred two thirds
less frequently than commits which mentioned \emph{Perl}, but
\emph{Perl} related topics appeared in the global static topics for
MaxDB while \emph{PHP} related topics did not.  Our local topic
analysis mentioned \emph{PHP} in 5 different topics, yet only
mentioned \emph{Perl} in 4 different topics. Perhaps the Perl mentions
were not as independent topics as the PHP topics were.

@@

\shrinkit
\subsection{Compare with topics over the entire development}
\shrinkit

@@

Previous work on topic analysis that employed LSI and LDA typically
extracted a certain number of topics and tracked them over time. This
means that a single, static set of topics was tracked across the whole
development history of a project. \emph{Global topic analysis} is
topic analysis applied to entire lifetime of a project.

We carried out the same kind of analysis, as shown in
Figure~\ref{fig:statictopics}. We extracted 20 topics and plotted the
number of messages per month that were related to that topic. We found
was that one topic would often dominate the results, as shown in the
third row of Figure \ref{fig:statictopics}, while messages related to
other topics did not appear very often.
% This technique seems useful if the 20 extracted topics are relevant
% and useful, that said one weakness of this approach is that documents
% for a topic might only occur in one time window, yet we are tracking
% messages that could be related to it over the entire lifetime of the
% project.
This approach seems reasonable if most of the extracted topics are of
broad interest during most of the development process.  However, it
may be that some topics are of strong interest but only briefly; in
such a case, a window-based model gives a much stronger indication of
the fleeting importance of such topics, and can help to put such a
topic into its proper context.

@@

If we approach the difference of global topic analysis and
windowed topic analysis via common tokens we can see that common token
tend to dominate both results, at least in the case of MaxDB 7.500 our
local topic approach netted us topics which contained these relatively
important and common words: \emph{UNICODE}, \emph{DEC/OSF1},
\emph{ODBC}, \emph{crash}, \emph{SQLCLI}, \emph{SYSDD.cpnix},
\emph{backup}, \emph{select}, \emph{make}, \emph{memory}, \emph{view},
and finally \emph{debug}. None of these tokens were part of the global 
topic analysis topics, but they were part of 566 commits (6\% of the
entire system) to MaxDB 7.500. These tokens were part of 87 out of 520
(26 months, 20 topics per month) of our locally generated topics.

@@

\begin{figure}
  \centering
  %\includegraphics[width=0.45\textwidth]{maxdb7500-everything-by-month}
  \includegraphics[width=0.46\textwidth]{month}
  \caption{MaxDB 7.500 topics analyzed with global topic analysis, 20
    topics (each row) and their document counts plotted over the
    entire development history of MaxDB 7.500 (26 months). The shade
    of the topic indicates the number of documents matching that topic
    in that month relative to the number of documents (white is most,
    black is least).}
  \label{fig:statictopics}
\end{figure}

@@

%   1. well i'm stumped
%   2. look at it?
%   3. investigate those revisions and check the proportion etc
%   4. exploratory
%   5.   work it out
%8.   Validity Threats [0/5]
\shrinkit
\section{Validity Threats}
\shrinkit

@@

\shrinkit
\section{Conclusions}
\shrinkit
%11.   Start file file:/home/abez/projects/lda-paper/lda-paper.tex

@@

We proposed and demonstrated the application of windowed topic
analysis, that is, which is topic analysis applied to commit messages during
periods of development. This approach differs from previous work that
would apply topic analysis globally to a project. We showed that many
topics that exist locally are relevant and interesting yet would often
not be detected via global topic analysis. We identify recurring topics
with a topic similarity measure which looks for topics which recur and
mutate repeatedly throughout the development of the software project.

The value of windowed topic analysis was demonstrated on a case study
of three database systems with a particular focus on MaxDB 7.500. We
showed that global topic analysis missed important topics like ODBC
changes.

@@

We presented several visualization techniques that focused on
different aspects of the data: temporality of trends, trend size, and
a compact-trend-view. The compact-trend-view shows much more
information than the views that global analysis could show, it
indicates how focused a period is by the total number of topics as
well shows topics by similarity so one can track trends across time.
Our trend-histogram immediately hi-lights and measures the size of
trends while our trend-time-line view shows how a topic reoccurs over
time. These visualizations help us get an over feeling about the
common and unique topics that developers focus on. If implemented
properly a user can easily zoom and query a topic.

\shrinkit
\subsection{ Future Work}
\shrinkit

One avenue of future work is automatic topic labelling. Given a word
distribution we should be able to automatically select a word or term
that describes that distribution. Potential external sources of topic
names include: software engineering taxonomies, ontologies and
standards.
